A molecular taxonomy of tumors independent of tissue-of-origin

### 2020-07-14 text/manuscript.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(rstan)
library(foreach)
library(sqldf)
library(tidyverse)
library(ggthemes)
library(data.table)
library(biomaRt)
library(dplyr)
library(GenomicRanges)
library(SummarizedExperiment)
library(reactome.db)
library(plyr)
library(cluster)
library(pheatmap)
library(clusterProfiler)
library(org.Hs.eg.db)
library(umap)
library(dbscan)
library(class)
library(TCGAbiolinks)
library(patchwork)
library(ggplotify)
library(ComplexHeatmap)
library(knitr)
library(kableExtra)
library(UpSetR)
library(cowplot)
```

@@

@@

@@

for(p in toppath){
  message(p)
  l_anno <- l_anno %>% mutate(p = 0) ## create a new col named p
  names(l_anno)[names(l_anno) == 'p'] <- p ## rename col with pathway name
  ## for any path_combo that includes the pathway name change the 0 into a 1
  if(p == "AUF1 (hnRNP D0) binds and destabilizes mRNA") { ## if it has this pathway grep a different way
    l_anno[grep(pattern = "AUF1", x = l_anno$path_combo), p] <- 1
  }
  if(p != "AUF1 (hnRNP D0) binds and destabilizes mRNA") { ## if it is not AUF1 do it normally
  l_anno[grep(pattern = p, x = l_anno$path_combo), p] <- 1 
  }
}
l_anno_ggplot <- list() ## list for storing ggplots
path_legend <- list()
for(n in 1:length(toppath)){
  message(n)
  p <- toppath[n]
  message(p)
  spec_path <- l_anno[,c("gene_name", p)]
  names(spec_path)[names(spec_path) == p] <- "p"
  path_ggplot <- ggplot(spec_path) +
  geom_bar(mapping = aes(x = 1, y = gene_name, fill = as.character(p)), 
           stat = "identity", 
           width = 1) +
  scale_fill_manual(values = r_anno_color, breaks = "1", name = "", labels = p) +
  theme_void() 
  path_legend[[n]] <- plot_grid(get_legend(path_ggplot), ncol = 1) ## extract legend from plot
  l_anno_ggplot[[n]] <- path_ggplot + theme(legend.position = "none") ## remove legend from plot
}

plot <- plot_grid(l_anno_ggplot[[1]], l_anno_ggplot[[2]], l_anno_ggplot[[3]], l_anno_ggplot[[4]], l_anno_ggplot[[5]], l_anno_ggplot[[6]], l_anno_ggplot[[7]], l_anno_ggplot[[8]], l_anno_ggplot[[9]], l_anno_ggplot[[10]], fig4, r_anno, align = "h", ncol = 12, axis = "tb", rel_widths = c(rep(0.1, 10), 15, 0.1))
plot

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rstan)
library(foreach)
library(sqldf)
library(tidyverse)
library(ggthemes)
library(data.table)
library(biomaRt)
library(dplyr)
library(GenomicRanges)
library(SummarizedExperiment)
library(reactome.db)
library(plyr)
library(cluster)
library(pheatmap)
library(clusterProfiler)
library(org.Hs.eg.db)
library(umap)
library(dbscan)
library(class)
library(TCGAbiolinks)
library(patchwork)
library(ggplotify)
library(ComplexHeatmap)
library(knitr)
library(kableExtra)
library(UpSetR)
library(cowplot)
```

@@

```{r mutation-burden-plot, message = FALSE}
umap_cancertype <- fread("../data/umap_2d_3d_coors.tsv")

mut_order <- list() ## store data in list 
for(n in 1:max(umap_cancertype$knn_clust)) { ## loop through all the clusters
  message(n)
  mut_order[[n]] <- umap_cancertype %>% filter(knn_clust == n) %>%  ## filter for specific class
    dplyr::select(V1, x_2d, y_2d, knn_clust, mut_count) %>% ## select relevant columns
    arrange(mut_count) ## arrange mutation in increasing order
}
mut_order <- mut_order %>% do.call(rbind, .) %>% as.data.frame() ## rbind list

mut_order$V1 <- factor(mut_order$V1, levels = mut_order$V1)
mut_order$knn_clust <- factor(mut_order$knn_clust, levels = c(1:12))

## Need the breaks positions for the x-axis
previous_class <- 0 ## starting position 1
position_of_previous_class <- 0 ## starting position 2
xaxis_breaks <- list() ## list for for loop
for(n in 1) { ## loop through 1 to 12 clusters
  message(n)
  class_spec <- round(mut_order %>% filter(knn_clust == n) %>% nrow(.) / 2, digits = 0) ## calculate halfway mark of class 
  xaxis_breaks[[n]] <- class_spec + previous_class + position_of_previous_class ## adding up all variables
  previous_class <- class_spec ## storing class halfway mark
  for(n in 2:max(as.numeric(as.character(mut_order$knn_clust)))) {
    message(n)
    position_of_previous_class <- xaxis_breaks[[n-1]] ## getting the previous class position
    class_spec <- round(mut_order %>% filter(knn_clust == n) %>% nrow(.) / 2, digits = 0) ## calculate halfway mark of new class
    xaxis_breaks[[n]] <- class_spec + previous_class + position_of_previous_class ## adding up all variables
    previous_class <- class_spec ## storing class halfway mark
  }
}
xaxis_breaks <- as.data.frame(do.call(rbind, xaxis_breaks)) ## rbind list
xaxis_breaks <- mut_order[rownames(mut_order) %in% xaxis_breaks$V1,] ## find the tumor samples in these positions

total_n <- as.data.frame(table(mut_order$knn_clust))

## for making dup discrete x-axis
guide_axis_label_trans <- function(label_trans = identity, ...) {
  axis_guide <- guide_axis(...)
  axis_guide$label_trans <- rlang::as_function(label_trans)
  class(axis_guide) <- c("guide_axis_trans", class(axis_guide))
  axis_guide
}

guide_train.guide_axis_trans <- function(x, ...) {
  trained <- NextMethod()
  trained$key$.label <- x$label_trans(trained$key$.label)
  trained
}

ggplot(mut_order, aes(x = V1, y = log10(mut_count + 1), color = knn_clust)) + 
  geom_point() + 
  ylab("Log10(Mutation Count)") + 
  xlab("") +
  scale_x_discrete(breaks = c(as.character(mut_order$V1[300]), as.character(xaxis_breaks$V1)), labels = c("Class =" ,1:12)) +
  guides(x.sec = guide_axis_label_trans(~paste(c("n = ", total_n$Freq)))) + ## create second x-axis
  scale_y_continuous(sec.axis = sec_axis(~.)) + ## dup y-axis
  #scale_y_log10() +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        legend.position = "none") +
  scale_color_manual(values = kelly.colours[-c(1:2)])
```

@@

class67_gene_2_path_heatmap$gene_name <- factor(class67_gene_2_path_heatmap$gene_name, levels = c(rev(gene_order)))

@@

for(p in toppath){
  message(p)
  l_anno <- l_anno %>% mutate(p = 0) ## create a new col named p
  names(l_anno)[names(l_anno) == 'p'] <- p ## rename col with pathway name
  ## for any path_combo that includes the pathway name change the 0 into a 1
  if(p == "AUF1 (hnRNP D0) binds and destabilizes mRNA") { ## if it has this pathway grep a different way
    l_anno[grep(pattern = "AUF1", x = l_anno$path_combo), p] <- 1
  }
  if(p != "AUF1 (hnRNP D0) binds and destabilizes mRNA") { ## if it is not AUF1 do it normally
  l_anno[grep(pattern = p, x = l_anno$path_combo), p] <- 1 
  }
}

l_anno_ggplot <- list() ## list for storing ggplots
path_legend <- list()

for(n in 1:length(toppath)){
  message(n)
  p <- toppath[n]
  message(p)
  spec_path <- l_anno[,c("gene_name", p)]
  names(spec_path)[names(spec_path) == p] <- "p"
  anno_color <- c("0" = "white", "1" = kelly.colours[2+n])
  path_ggplot <- ggplot(spec_path) +
  geom_bar(mapping = aes(x = 1, y = gene_name, fill = as.character(p)), 
           stat = "identity", 
           width = 1) +
  scale_fill_manual(values = anno_color, breaks = "1", name = "", labels = p) +
  theme_void() 
  path_legend[[n]] <- plot_grid(get_legend(path_ggplot), ncol = 1) ## extract legend from plot
  l_anno_ggplot[[n]] <- path_ggplot + theme(legend.position = "none") ## remove legend from plot
}

plot <- plot_grid(l_anno_ggplot[[1]], l_anno_ggplot[[2]], l_anno_ggplot[[3]], l_anno_ggplot[[4]], l_anno_ggplot[[5]], l_anno_ggplot[[6]], l_anno_ggplot[[7]], l_anno_ggplot[[8]], l_anno_ggplot[[9]], l_anno_ggplot[[10]], fig4, r_anno, align = "h", ncol = 12, axis = "tb", rel_widths = c(rep(0.1, 10), 15, 0.1))
plot_legend <- plot_grid(path_legend[[1]], path_legend[[2]], path_legend[[3]], path_legend[[4]], path_legend[[5]], path_legend[[6]], path_legend[[7]], path_legend[[8]], path_legend[[9]], path_legend[[10]], ncol = 5)

plot_grid(plot, plot_legend, ncol = 1, rel_heights = c(10, 1.5))
